---
title: "Monthly Climate Data Organization"
author: "Anna Talucci"
date: "June 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview

GEE data 

Climate data from Terra Climate 

Temperature (degrees celcius) scale = 0.1
precipitation (mm) no scaling

Monthly data

# Packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(readr)
library(cowplot)
library(ggcorrplot)
```



# Data

## Monthly

```{r}
month_path <- "../data/climate_monthly"   # path to the data
month_files <- dir(month_path, pattern = "*.csv") # get file names
month_files
```

```{r}
month_data <- month_files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(month_path, .), .id = "file_name")) %>% 
  reduce(rbind)
month_data
```
map_df(~read_csv(.x, col_types = cols(), col_names = FALSE), .id = "file_name")


```{r}
list_of_files <- list.files(path = "../data/climate_monthly",
                            full.names = TRUE)
list_of_files
df <- list_of_files %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x, col_types = cols(), col_names = TRUE), .id = "file_name")     
df
```
 separate(file_name, c("A","B","C"), sep = "([.?:])")
```{r}
df1 = df %>% 
  separate(file_name, c("A","B","C","D","E","F","G"), sep = "([..///_])", extra = "merge") %>%
  separate(G, c("A1","B1","C1","D1","E1","F1","G1"), sep = "([._])", extra = "merge") %>%
  dplyr::select(A1:B1, ID_obj:vs) %>%
  rename(climate_yr = A1, climate_mon = B1)

df1
  
```




```{r}
win_temp_data %>% summarise_all(~ sum(is.na(.)))
```

```{r}
win_temp_sel = win_temp_data %>% 
                      dplyr::select(ID_obj:mean) %>%
                      rename(wtr_temp = mean)
                      

head(win_temp_sel)
```


## Snow


```{r}
snow_path <- "../data/gee_perimeter/snow_off"   # path to the data
snow_files <- dir(snow_path, pattern = "*.csv") # get file names
```

```{r}
snow_data <- snow_files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(snow_path, .))) %>% 
  reduce(rbind)
snow_data
```

```{r}
snow_data %>% summarise_all(~ sum(is.na(.)))
```

```{r}
snow_sel = snow_data %>% 
                      dplyr::select(ID_obj:year) 
head(snow_sel)
```

## Combine Temp Precip and snow

```{r}
df_climate = merge(df_temp_pr, snow_sel, by = c("ID_obj", "area_ha", "area_m", "fire_yr"))

```

```{r}
df_climate
```

```{r}
df_climate %>% summarise_all(~ sum(is.na(.)))
```

```{r}
na_data <- df_climate %>% filter_all(any_vars(is.na(.))) 
na_data
```

# Clean data
- Remove NA
- Scale temperature data
```{r}
head(df1)
```

```{r}
df1 %>% 
  na.omit() %>%
  mutate(tmmx_scale = 0.1) %>%
  mutate(tmp_mx = tmmx*tmmx_scale) %>%
  mutate(fire_yr = as.factor(fire_yr))



```

```{r eval=FALSE, include=FALSE}
write.csv(df_climate1,"../outputs/fire_climate_data.csv", row.names = FALSE)
```

# Summarize data

```{r}
df_sum = df_climate1 %>% group_by(fire_yr) %>%
  select(area_ha, wtr_temp, spr_temp, smr_temp, wtr_pr, spr_pr, smr_pr, calDoy) %>% # select variables to summarise
  summarise_each(funs(min = min, 
                      q25 = quantile(., 0.25), 
                      median = median, 
                      q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

head(df_sum)
```

# Define graph Theme and color Palette 

```{r}
boxplot_theme = theme_bw() + theme(legend.position = "none") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))
```
theme(panel.grid = element_blank(),
        panel.border = element_blank())
```{r}
reg_theme = theme_bw() + theme(legend.position = "none") +
  theme(plot.margin = unit(c(t = 0.3, r = 0.3, b = 0.3, l = 0.1), "cm")) +
  theme(axis.title.x = element_text(size = 10, hjust = 0.5, vjust = -0.2),
        axis.title.y = element_text(size = 10, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 9, color = "black"),
        axis.text.y = element_text(size = 9, color = "black"),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))
```



# Visualize data

## Boxplots all varaibles

```{r}
plot1 = ggplot(df_climate1, aes(x = fire_yr, y = wtr_temp)) +
  geom_boxplot() +
  boxplot_theme

plot1
```

```{r}
plot2 = ggplot(df_climate1, aes(x = fire_yr, y = spr_temp)) +
  geom_boxplot() +
  boxplot_theme

plot2 
```

```{r}
plot3 = ggplot(df_climate1, aes(x = fire_yr, y = smr_temp)) +
  geom_boxplot() +
  boxplot_theme

plot3
```

```{r}
plot4 = ggplot(df_climate1, aes(x = fire_yr, y = wtr_pr)) +
  geom_boxplot() +
  boxplot_theme

plot4
```

```{r}
plot5 = ggplot(df_climate1, aes(x = fire_yr, y = spr_pr)) +
  geom_boxplot() +
  boxplot_theme
plot5
```

```{r}
plot6 = ggplot(df_climate1, aes(x = fire_yr, y = smr_pr)) +
  geom_boxplot() +
  boxplot_theme
plot6
```

```{r}
plot7 = ggplot(df_climate1, aes(x = fire_yr, y = calDoy)) +
  geom_boxplot() +
  boxplot_theme
plot7
```



## Bar plot of means 
```{r}
head(df_sum)
```

 
## Winter

```{r}
ggplot(df_sum, aes(x = fire_yr, y = wtr_temp_mean)) +
  geom_bar(stat="identity", color="black") +
  geom_errorbar(aes(ymin=wtr_temp_mean-wtr_temp_sd, ymax=wtr_temp_mean+wtr_temp_sd), width=.2) +
  boxplot_theme
```   



```{r}
ggplot(df_climate1, aes(x = fire_yr, y = wtr_temp)) +
  geom_bar(stat = "summary_bin", fun = mean) +
  boxplot_theme
```
## Spring Temp
```{r}
ggplot(df_sum, aes(x = fire_yr, y = spr_temp_mean)) +
  geom_bar(stat="identity", color="black") +
  geom_errorbar(aes(ymin=spr_temp_mean-spr_temp_sd, ymax=spr_temp_mean+spr_temp_sd), width=.2) +
  boxplot_theme
```   
```{r}
ggplot(df_climate1, aes(x = fire_yr, y = spr_temp)) +
  geom_bar(stat = "summary_bin", fun = mean) +
  boxplot_theme
```
## SUmmer Temp
```{r}
ggplot(df_sum, aes(x = fire_yr, y = smr_temp_mean)) +
  geom_bar(stat="identity", color="black") +
  geom_errorbar(aes(ymin=smr_temp_mean-smr_temp_sd, ymax=smr_temp_mean+smr_temp_sd), width=.2) +
  boxplot_theme
```   

# Coorelations
```{r}
head(df_climate1)
```


```{r}
df_climate1_corr = df_climate1 %>%
  dplyr::select(area_ha, wtr_temp:calDoy, year)

corr = round(cor(df_climate1_corr), 1)  
p_mat <- round(cor_pmat(df_climate1_corr), 3)
p_mat
```

```{r fig.height=6, fig.width=6}
cor_plot = ggcorrplot(corr)
cor_plot
```
```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-01_CorrPlot-Temp-PR-Snowoff.jpeg", plot = cor_plot, width = 6, height = 6, units = c("in"), dpi=600 )
```

# Combine plots

```{r fig.height=11, fig.width=12}
vis1 = cowplot::plot_grid(plot1, plot4, plot2, plot5, plot3, plot6, plot7,  labels = "AUTO", align = "hv", label_size = 12, ncol= 2) + 
  theme(plot.margin = margin(0, 0, 0, 0)) 

vis1
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-01_Boxplot-Temp-PR-Snowoff.jpeg", plot = vis1, width = 12, height = 11, units = c("in"), dpi=600 )
```