---
title: "Monthly Climate Data Organization"
author: "Anna Talucci"
date: "June 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview

GEE data 

Climate data from Terra Climate 

Temperature (degrees celcius) scale = 0.1
precipitation (mm) no scaling

Monthly data


***
# Data Details
[Terra Climate](https://developers.google.com/earth-engine/datasets/catalog/IDAHO_EPSCOR_TERRACLIMATE)

Name  |	Units	| Scale	| Description
------|-------|-------|------------
aet	  | mm	  |	0.1	  | Actual evapotranspiration,  
def	  | mm	  |	0.1	  | Climate water deficit, 
pdsi	|       |	0.01	| Palmer Drought Severity Index
pet	  | mm	  |	0.1	  | Reference evapotranspiration (ASCE Penman-Montieth)
pr	  | mm		|	      | Precipitation accumulation
ro	  | mm	  |       |	Runoff, derived using a one-dimensional soil water balance model
soil	| mm	  | 0.1	  | Soil moisture, derived using a one-dimensional soil water balance model
srad	| W/m^2	|	0.1	  | Downward surface shortwave radiation
swe	  | mm	  |		    | Snow water equivalent, 
tmmn	| °C	  |	0.1	  | Minimum temperature
tmmx	| °C	  |	0.1	  | Maximum temperature
vap	  | kPa	  |	0.001	| Vapor pressure
vpd   |	kPa	  |	0.01	| Vapor pressure deficit
vs	  | m/s	  |	0.01	| Wind-speed at 10m

***

# Packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(readr)
library(cowplot)
library(ggcorrplot)
```

# Define graph Theme and color Palette 

```{r}
boxplot_theme = theme_bw() + theme(legend.position = "none") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))
```
theme(panel.grid = element_blank(),
        panel.border = element_blank())
```{r}
reg_theme = theme_bw() + theme(legend.position = "none") +
  theme(plot.margin = unit(c(t = 0.3, r = 0.3, b = 0.3, l = 0.1), "cm")) +
  theme(axis.title.x = element_text(size = 10, hjust = 0.5, vjust = -0.2),
        axis.title.y = element_text(size = 10, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 9, color = "black"),
        axis.text.y = element_text(size = 9, color = "black"),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))
```



# Data

## Monthly

```{r}
month_path <- "../data/climate_monthly"   # path to the data
month_files <- dir(month_path, pattern = "*.csv") # get file names
month_files
```

```{r}
month_data <- month_files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(month_path, .), .id = "file_name")) %>% 
  reduce(rbind)
month_data
```
map_df(~read_csv(.x, col_types = cols(), col_names = FALSE), .id = "file_name")


```{r}
list_of_files <- list.files(path = "../data/climate_monthly",
                            full.names = TRUE)
list_of_files
df <- list_of_files %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x, col_types = cols(), col_names = TRUE), .id = "file_name")     
df
```
 separate(file_name, c("A","B","C"), sep = "([.?:])")
```{r}
df1 = df %>% 
  separate(file_name, c("A","B","C","D","E","F","G"), sep = "([..///_])", extra = "merge") %>%
  separate(G, c("A1","B1","C1","D1","E1","F1","G1"), sep = "([._])", extra = "merge") %>%
  dplyr::select(A1:B1, ID_obj:vs) %>%
  rename(climate_yr = A1, climate_mon = B1)

df1
  
```





## Snow


```{r}
snow_path <- "../data/gee_perimeter/snow_off"   # path to the data
snow_files <- dir(snow_path, pattern = "*.csv") # get file names
```

```{r}
snow_data <- snow_files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(snow_path, .))) %>% 
  reduce(rbind)
snow_data
```

```{r}
snow_data %>% summarise_all(~ sum(is.na(.)))
```

```{r}
snow_sel = snow_data %>% 
                      dplyr::select(ID_obj:year) 
head(snow_sel)
```

## Combine Temp Precip and snow

```{r}
df_climate = merge(df_temp_pr, snow_sel, by = c("ID_obj", "area_ha", "area_m", "fire_yr"))

```

```{r}
df_climate
```

```{r}
df_climate %>% summarise_all(~ sum(is.na(.)))
```

```{r}
na_data <- df_climate %>% filter_all(any_vars(is.na(.))) 
na_data
```

# Clean data
- Remove NA
- Scale temperature data
```{r}
df1
```

```{r}
df2 = df1 %>% 
  na.omit() %>%
  mutate(tmmx= tmmx*0.1) %>%
  mutate(def = def*0.1) %>%
  mutate(fire_yr = as.factor(fire_yr))


df2
```



# Summarize data

```{r}
df_sum = df_climate1 %>% group_by(fire_yr) %>%
  select(area_ha, wtr_temp, spr_temp, smr_temp, wtr_pr, spr_pr, smr_pr, calDoy) %>% # select variables to summarise
  summarise_each(funs(min = min, 
                      q25 = quantile(., 0.25), 
                      median = median, 
                      q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

head(df_sum)
```



# Visualize data

## Boxplots all varaibles

```{r}
plot1 = ggplot(df2, aes(x = fire_yr, y = tmp_mx, color=climate_mon)) +
  geom_boxplot() +
  theme_bw()

plot1
```

```{r}
plot2 = ggplot(df2, aes(x = fire_yr, y = pr, color=climate_mon)) +
  geom_boxplot() +
  theme_bw()

plot2
```

 

# Coorelations
```{r}
head(df2)

```

```{r}
qplot(data = df2, x=climate_yr, y=def, color = climate_mon)
```


```{r}
qplot(data = df2, x=climate_yr, y=tmmx, color = climate_mon)
```

```{r}
qplot(data = df2, x=climate_yr, y=pr, color = climate_mon)
```
```{r}
head(df2)
```

```{r}
df3 = df2 %>%
  dplyr::select(climate_mon:area_m, def, tmmx, pr)
df3
```

mutate(gradebook, letter = ifelse(grade %in% 60:69, "D",
                                     ifelse(grade %in% 70:79, "C",
                                            ifelse(grade %in% 80:89, "B",
                                                   ifelse(grade %in% 90:99, "A", "F")))))


```{r}
df3 %>% pivot_wider(names_from = climate_mon, values_from = c(aet, def, tmmx, pr))
```






```{r}
df_climate1_corr = df2 %>%
  mutate(year = as.numeric(climate_yr)) %>%
  dplyr::select(area_ha, tmp_mx, pr, year)

str(df_climate1_corr,)
```

```{r}
corr = round(cor(df_climate1_corr), 1)  
p_mat <- round(cor_pmat(df_climate1_corr), 3)
p_mat
```

```{r fig.height=6, fig.width=6}
cor_plot = ggcorrplot(corr)
cor_plot
```
```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-01_CorrPlot-Temp-PR-Snowoff.jpeg", plot = cor_plot, width = 6, height = 6, units = c("in"), dpi=600 )
```

# Combine plots

```{r fig.height=11, fig.width=12}
vis1 = cowplot::plot_grid(plot1, plot4, plot2, plot5, plot3, plot6, plot7,  labels = "AUTO", align = "hv", label_size = 12, ncol= 2) + 
  theme(plot.margin = margin(0, 0, 0, 0)) 

vis1
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-01_Boxplot-Temp-PR-Snowoff.jpeg", plot = vis1, width = 12, height = 11, units = c("in"), dpi=600 )
```