---
title: "Data Organization"
author: "Anna Talucci"
date: "June 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# overview



# Packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(readr)
```



# Data

## Temerature

### Winter

### Spring
files <- list.files(path = "../data/gee_perimeter/spring_temp", pattern = "*.csv")
files
```{r}


data_path <- "../data/gee_perimeter/spring_temp"   # path to the data
files <- dir(data_path, pattern = "*.csv") # get file names
```

```{r}
data <- files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(data_path, .))) %>% 
  reduce(rbind)
data
```


```{r eval=FALSE, include=FALSE}
spt2001= read.csv("../data/gee_perimeter/2001_fires_spring_temp.csv")
spt2002= read.csv("../data/gee_perimeter/2002_fires_spring_temp.csv")
spt2003= read.csv("../data/gee_perimeter/2003_fires_spring_temp.csv")
spt2004= read.csv("../data/gee_perimeter/2004_fires_spring_temp.csv")
spt2005= read.csv("../data/gee_perimeter/2005_fires_spring_temp.csv")
spt2006= read.csv("../data/gee_perimeter/2006_fires_spring_temp.csv")
spt2007= read.csv("../data/gee_perimeter/2007_fires_spring_temp.csv")
spt2008= read.csv("../data/gee_perimeter/2001_fires_spring_temp.csv")
spt2009= read.csv("../data/gee_perimeter/2001_fires_spring_temp.csv")
spt2010= read.csv("../data/gee_perimeter/2001_fires_spring_temp.csv")
spt2011= read.csv("../data/gee_perimeter/2001_fires_spring_temp.csv")
```


## Precipitation


## Snow